Fused regression for multi-source gene regulatory network inference

### 2016-07-30 paper_revisions.tex
zack changes. Figure legends may need to be changed more if the layout changes

\subsection{Simulations}
\subsubsection{Using fused regression to learn related networks}
We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions. Simulated networks consisted of pairs of related 10 TF by 200 gene networks, with varying numbers of expression data samples made available to the solver. In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network. When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (figure \ref{fusion}b). Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.

@@

\subsubsection{Adaptive fusion successfully identifies and unfuses 'neofunctionalized' genes}
Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. Some gene interactions may be non-conserved, and in order to minimize bias it would be desirable to identify these interactions and relax the fusion constraints on them. To allow for orthologous genes to be unfused we implemented an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (figure \ref{adapt-schematic}). 
Pairs of interactions that are dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow. 
Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data. 
As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints is much more direct evidence for neofunctionalization than comparing separately fit networks could provide.

We performed a simulation to assess the ability of our adaptive fusion algorithm to distinguish which parts of two input networks are conserved vs. neofunctionalized (figure \ref{adaptivefusion}). We generated synthetic fused networks and introduced error in the fusion constraints by adding false positives and negatives to the orthology information given to the solver. 
Because we knew which entries in the orthology mapping were ``incorrect'' (not reflected in the generation of the networks), we could correctly label fusion constraints that involved one or more ``incorrect'' mappings. 
We verified that adaptive-fusion unfused mostly ``incorrectly fused'' interactions (figure \ref{adaptivefusion}a red dots), while leaving truly analogous interactions fused (figure \ref{adaptivefusion}a green dots).

We then compared the interaction weights which were accurately fused, and interaction weights which were inaccurately fused due to incorrect orthology information (figure \ref{adaptivefusion}b). Because fused L2 heavily penalizes large differences between weights which are predicted to be similar, all learned interaction weights with fusion constraints are similar. Those constraints that are unfused by adaptive fusion are allowed to differ; some of the incorrectly fused constraints have large differences in interaction weights.

@@

To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B. subtilis} interactions with corresponding activation and repression sign. 
Of these 3,040 priors, 968 had corresponding interactions in \textit{B. anthracis}. 
Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species. 
However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. 
As a result, we can only evaluate network recovery for \textit{B. subtilis}, and \textit{B. subtilis} also has the majority of our conditions. 
In order to simulate the data-poor regime, we subsampled our \textit{B. subtilis} data. 
We divided our \textit{B. subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B. subtilis} data from that fold alone fused to the entire 72 \textit{B. anthracis} conditions (figure \ref{xspecies-real}a). 
Though overall performance is hindered by our subsampling of \textit{B. subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B. subtilis} network when using fused regression (figure \ref{xspecies-real}a). 
Notably, these performance gains occur mostly at low values of recall. That is: the highest confidence part of the network is inferred more accurately, with minimal gains for interactions which are more uncertain. Because these interactions are likely to be the focus of followup experiments and validation, gains here are more valuable for prioritizing the order in which interactions are investigated.

\subsubsection{Adaptive fusion using bacterial data}
[ZACK MORE HERE TOO]
 The goal of adaptive fusion is to unfuse constraints between non-conserved interactions, while leaving intact all other constraints. 
 However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B. subtilis} and \textit{B. anthracis}, we were unable to directly evaluate how accurately adaptive fusion identified these interactions.
  We opted instead to introduce a large number of random fusion constraints between genes not known to be orthologous. 
 These fake constraints, which are unlikely to reflect any conserved network structure, served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
 We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints, along with those generated by known orthology. 
 We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see figure \ref{xspecies-real}b & c). 
 Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (figure \ref{xspecies-real}c).

@@

Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
We included, along with our \textit{B. subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B. subtilis} \cite{nicolas2012condition}. 
We compared performance when learning the networks separately and then rank combining (as in \cite{ciofani_validated_2012}) to learning the networks simultaneously using fusion regression, and we show improvement in performance using our fused L2 approach (figure \ref{fusedl2-real}a).

Our approach can also make use of forms of biological data other than RNA expression. Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts. In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}. We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors. We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon. We also used a subset of the gold-standard as priors by relaxing the $\lambda_R$ parameter for these interactions, as in \cite{greenfield_robust_2013}, and evaluated on the remaining set of priors. Incorporation of operon-derived fusion constraints in \textit{B. subtilis} improved network recovery (figure \ref{fusedl2-real}b).

\subsection{Transcription factor activity estimation integrates into fusion regression approach}
We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA). 
Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as applied to \textit{B. subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}. 
TFA activity estimates  transcription factor activities that are modulated through mechanisms such as dimerization and interaction with required factors.
TFA activity estimates have been shown prior to be better predictors of TF function than expression level alone in several contexts including similar network inference tasks\cite{fu_reconstructing_2011} \cite{arrieta-ortiz_experimentally_2015}. 
We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
To test the integration of this approach with our fused regression, we assessed the combination of \textit{B. subtilis} datasets, as in figure \ref{fusedl2-real}a, with the incorporation of TFA estimation. 
We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
The remaining interactions were reserved as a gold standard for validation. 
As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (figure \ref{tfa}). We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.

@@

This is equivalent to the original problem due to the block structure of matrix multiplication. 
In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary. 
However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (figure \ref{cartoon}). 
As a result, these linked response variables must be solved simultaneously through vectorization. 
Two response variables are linked if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them. 
However, every other response variable can be solved separately. 
In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints. 
If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractable.

@@

And for steady state conditions
\begin{equation}
\alpha_{i}x_{i}(t_k) = \sum \beta_{i,j}x_{j}(t_k)
\end{equation}
Both kinds of conditions can be included in the set of equations to be solved. We can summarize these equations in matrix form as
\begin{equation}
Y = X \beta 
\end{equation}
where $Y$ is the gene expression matrix, $X$ is the TF expression matrix, and $\beta$ is the regulatory weights we are interested in learning.
We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene. 
Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction. The matrix $\beta$ can be solved using linear regression. 
Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
\begin{equation}
\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
\end{equation}
This is similar to the formulation used in the \textit{Inferelator} algorithm, except that we use an $L2$ penalty in place of the more standard $L1$ penalty. Although $L1$ penalties have the theoretical appeal of sparsity - which matches our assumption of sparsity in biological networks - we are primarily interested here in ranking interactions, rather than recovering a specific network. Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{}. Although our approach can be easily modified to use an $L1$ penalty, we did not observe performance gains when doing so, and the version presented here is in some ways conceptually simpler (with the combined penalty function being equivalent to a multivariate Gaussian prior).

@@

\textit{B. subtilis} and \textit{B. anthracis} are distantly related bacterial species with limited gene orthology. Nevertheless we show that network recovery in \textit{B. subtilis} can be improved through the inclusion of expression data from \textif{B. anthracis}. Many previous methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}. This assumption may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (62\% and 67\%) of the \textit{B. subtilis} and \textit{B. anthracis} genomes do not have clear orthologs. Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.

@@

\pagebreak
\begin{figure}[h!]
    \centering
    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. 
    Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes. 
    Rectangles represent matrices; because weights can be solved independently unless there exists fusion constraints between them, we identify related weights and construct matrix for solving. }
\end{figure}
\pagebreak

\begin{figure}
\begin{center}
  \includegraphics[scale=0.45]{fig1.pdf}
  \caption{\label{fusion} \textbf{A.} Schematic representation of the the generation of fusion constraints from orthology mappings. Dashed arrows indicate potential regulatory interactions, while solid arrows denote orthology. We introduce fusion constraints for pairs of interactions for which both the regulator and regulated gene are orthologs of one another. 
  In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions. \textbf{B.} 
  In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity). Mean squared error of the inferred vs. true coefficient matrices for network 1 are plotted as a function of the number of conditions generated for species 1 (x-axis) and the number of conditions generated for species 2 (y-axis). 
  As expected, increasing the number of samples available for the species of interest improves network inference performance. 
  However, because we are fusing to data from a related species, similar gains are observed when increasing the amount of data available in this second species. 
  \textbf{C-F} Show the varying effects of fusion on simulated networks with different levels of conservation. We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks (Amount fused, top). For each network, we evaluated AUPR on one of the species for: all interactions (blue line), interactions with fusion constraints (green line), and interactions without fusion constraints (orange line). At every level of conservation, constrained interactions show the largest benefits of fusion, with the magnitude of the benefit growing with fraction of orthologous genes. 
  When the networks are highly conserved, however, even interactions that are not directly constrained through fusion are recovered more accurately as $\lambda_S$ increases.}
\end{center}
\end{figure}
\pagebreak

@@

\begin{figure}
\begin{center}
  \includegraphics[scale=0.45]{fig3.pdf}
  \caption{\label{adapt-schematic} Adaptive fusion loss function (\textbf{A}) and derivative of loss function (\textbf{B}). 
  \textbf{A.} Adaptive fusion is a quadratic around the origin, begins to taper at $a/2$, and plateaus at $a$. After the plateau, increasing the difference in interaction weight of fused interactions does not further affect the penalty incurred through fusion. As a result, interaction weights in this zone are effectively unfused from one another (the fusion penalty behaves like a constant). 
  \textbf{B.} Shows the derivative of the adaptive fusion penalty, which is used to implement adaptive fusion through local quadratic approximation. The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions and like these penalties has a zero derivative far from the origin. }
\end{center}
\end{figure}
\pagebreak

\begin{figure}
\begin{center}
  \includegraphics[scale=0.45]{fig4_2.pdf}
  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in unfusing interactions that have diverged through evolution, we performed a series of simulations inferring networks given a partially corrupted list of orthology mappings. 
  Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage and 40\% false orthology coverage. \textbf{A.} We plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$). 
  Interactions that are generated from false orthologs are marked as red, while interactions generated from true orthologs are shown in green. 
  As expected, false fusion constraints give rise to uncorrelated weights (red dots). \textbf{B.} 
  When fit with fused-L2, fusion constraints give rise to very similar weights in the two species for 'true' and 'false' interactions. 
  \textbf{C.} Adaptive fusion run on the same network unfuses constraints for which the inferred weights are dissimilar beyond a certain point. 
  Here, unfused interactions are almost entirely interactions between false orthologs. \textbf{D-F} Show the distribution of the absolute value of the difference in inferred weight for interactions with true fusion constraints (green) and false fusion constraints (orange)}
\end{center}
\end{figure}

>>>>

\subsection{Simulations}
\subsubsection{Using fused regression to learn related networks}
We created synthetic networks to approximate two related biological processes, then evaluated performance of our fused L2 regression, which learns the networks simultaneously given a prior on the relatedness of interactions. Simulated networks consisted of pairs of related 10 TF by 200 gene networks, with varying numbers of expression data samples made available to the solver. In order to compare the relative contribution of data in the species of interest (species 1) and in a related organism (species 2) to network recovery, we varied the number of conditions in each species while evaluating the mean-squared-error between the recovered network for species 1 and its known true network. When the amount of data from the related species was held constant, increasing the amount of data available for learning the network for the species of interest resulted in a more accurate network prediction, as expected  (Figure \ref{fusion}b). Similarly, when we increased the amount of data from the related species, we obtained performance gains on network one using fused L2 regression, demonstrating our ability to improve network inference on one dataset through incorporation of a related dataset.

@@

\subsubsection{Adaptive fusion with simulated data}
Although we expect that orthology provides some evidence of regulatory similarity, we know that it is not a perfect proxy for functional conservation \cite{gabaldon_functional_2013, studer_how_2009, nehrt_testing_2011}. Some gene interactions may be non-conserved, and in order to minimize bias it would be desirable to identify these interactions and relax the fusion constraints on them. To this end, we developed an adaptive fusion algorithm that attempts to optimize a nonconvex saturating penalty function on differences between fused interactions (Figure \ref{adapt-schematic}). Pairs of interactions that are dissimilar even after fusion, which sit in the flat portion of this penalty function, are effectively ``unfused,'' and no further penalty is incurred as differences in interaction weights grow. Our network procedure strongly favors similarity of fused interactions, and only ``unfuses'' interactions when their similarity cannot be reconciled with expression data. As a result, the ``unfusing'' or relaxation of the fusion penalty on certain constraints serves not only to reduce bias in the inferred network weights, but can be interpreted directly as evidence against conservation.

We performed a simulation to assess the performance of adaptive fusion in the context of partially conserved networks. We generated synthetic fused networks based on a partial orthology mapping, then randomly assigned the remaining genes unrelated orthologs. These false orthologs gave rise to false fusion constraints, which if enforced would degrade network inference performance. Slightlymore than half of the fusion constraints in the generated network were false. We were interested in the ability of adaptive fusion to identify these false fusion constraints, as well as improve the accuracy of the inferred networks by reducing the bias associated with fusion. We verified that solving these partially conserved networks with fused-L2 forced interactions bound by both true and false fusion constraints to be similar (Figure \ref{adaptivefusion}b). Fusion led to improved performance on the conserved part of the network and degraded performance on the non-conserved part of the network (Figure \ref{adaptivefusion}d, Fused-L2). Applying adaptive fusion with the ``a'' parameter set to a percentile that roughly matched the fraction of false fusion constraints preserved the performance gains obtained through fusion on the conserved part of the network, while substantially reducing the loss on the non-conserved part of the network. Many of the incorrectly fused interactions were relaxed, allowing these interactions to freely fit the data without bias from a faulty prior. These non-conserved interactions -- which have diverged from one another and left the diagonal in  Figure \ref{adaptivefusion}c -- can be identified through inspection of the $\lambda_S$ parameter following adaptive fusion, which falls to zero for unfused interactions. This application suggests that adaptive fusion has use as both a tool to reduce bias while retaining many of the benefits of fused regression, and as a method for determining conservation across networks.

@@

To assess network inference performance, and for use as priors, we used a gold standard of 3,040 known \textit{B. subtilis} interactions with corresponding activation and repression sign. 
Of these 3,040 priors, 968 had corresponding interactions in \textit{B. anthracis}. 
Based on our simulation results, we can expect the greatest gains in network-inference performance from fusion when the species of interest has a small number of available conditions, but data is abundant in a related species. 
However, in order to evaluate performance objectively a gold-standard of known interactions is necessary. 
As a result, we can only evaluate network recovery for \textit{B. subtilis}, and \textit{B. subtilis} also has the majority of our conditions. 
In order to simulate the data-poor regime, we subsampled our \textit{B. subtilis} data. 
We divided our \textit{B. subtilis} data into $k$ folds, and then for each fold fit a network to the \textit{B. subtilis} data from that fold alone fused to the entire 72 \textit{B. anthracis} conditions (Figure \ref{xspecies-real}a). 
Though overall performance is hindered by our subsampling of \textit{B. subtilis} data (a necessary procedure to allow evaluation of networks) we demonstrate marked improvement in learning the \textit{B. subtilis} network when using fused regression (Figure \ref{xspecies-real}a). 
Notably, these performance gains occur mostly at low values of recall. That is: the highest confidence part of the network is inferred more accurately, with minimal gains for interactions which are more uncertain. Because these interactions are likely to be the focus of followup experiments and validation, gains here are more valuable for prioritizing the order in which interactions are investigated.

\subsubsection{Adaptive fusion using bacterial data}
When we applied adaptive fusion to our bacterial datasets we did not see a significant improvement in network recovery relative to fused-L2. Based on simulation results, we expected that non-conserved interactions weights would be fit more accurately with adaptive-fusion than with fused-L2, but that both algorithms would perform worse than independently fitting the networks on this subset of interactions. It may be that the quality of the independently fit network was too poor in this case for fusion constraints on non-conserved interactions to meaningfully degrade network inference performance. Nevertheless, we were interested in the application of adaptive fusion to identifying neo-functionalized or non-conserved interactions. Essentially, we wanted to identify constraints arising from orthology that does not correspond to functional conservation. However, because we lacked a comprehensive gold standard of known non-conserved interactions between \textit{B. subtilis} and \textit{B. anthracis}, we were forced to indirectly evaluate the performance of adaptive fusion in identifying these interactions.

We obtained a known set of non-conserved interactions by randomly generating orthology mappings between genes not known to be orthologous. These orthology mappings generated false constraints, as in our simulation studies,  which are unlikely to reflect any conserved network structure. We used these constraints served as a proxy for the unknown fraction of non-conserved interactions between orthologous genes. 
 We ran adaptive fusion to learn networks for \textit{B. subtilis} and \textit{B. anthracis}, using these constraints along with those generated by known orthology. We confirmed that our injected spurious fusion constraints were unfused at a higher rate than those generated by known orthologs (see Figure \ref{xspecies-real}b & c). Although it may seem odd that a large fraction of fake constraints were left intact, we note that biological networks tend to be sparse, so that many of the random fusion constraints are between coefficients with near zero weight (and therefore near zero difference in weight) (Figure \ref{xspecies-real}c). It is unclear to what extent the relatively large fraction of fusion constraints between true orthologs that were unfused reflects an inadequacy of the method, or a relatively high rate of neofunctionalizaton in these organisms.

@@

Currently, the most widely used approach to combining datasets for network inference is to learn networks from disparate datasets separately, then rank combine the networks as in Marbach et al \cite{marbach_revealing_2010}.
We included, along with our main \textit{B. subtilis} dataset, a previously published dataset containing 269 samples covering 104 conditions, obtained using a different tiling microarray (vs custom microarray) and different strain of \textit{B. subtilis} \cite{nicolas2012condition}. 
We compared performance when learning the networks separately and then rank combining (as in \cite{ciofani_validated_2012}) to learning the networks simultaneously using fusion regression, and we show improvement in performance using our fused L2 approach (Figure \ref{fusedl2-real}a).

Our approach can also make use of forms of biological data other than RNA expression. Information about the expected similarity of TF-gene interactions can come from knowledge about the promoter region or the structure, for bacteria, of polycistronic transcripts. In bacteria, genes within the same operon are typically under the control of the same promoter \cite{lawrence_shared_2002}. We posited, therefore, that genes within the same operon will be regulated similarly by the same transcription factors. We applied fusion regression by creating fusion constraints between a given transcription factor and genes within the same operon. We also used a subset of the gold-standard as priors by relaxing the $\lambda_R$ parameter for these interactions, as in \cite{greenfield_robust_2013}, and evaluated on the remaining set of priors. Incorporation of operon-derived fusion constraints in \textit{B. subtilis} improved network recovery (Figure \ref{fusedl2-real}b).

\subsection{Transcription factor activity estimation integrates into fusion regression approach}
We tested a combination of our fused regression approach with a method for estimating transcription factor activities (TFA). 
Rather than modeling gene expression using transcription factor mRNA abundance, we fit gene expression as a function of transcription factor activity, as applied to \textit{B. subtilis} by Arrieta-Ortiz et al \cite{arrieta-ortiz_experimentally_2015}. 
TFA activity estimates  transcription factor activities that are modulated through mechanisms such as dimerization and interaction with required factors.
TFA activity estimates have been shown prior to be better predictors of TF function than expression level alone in several contexts including similar network inference tasks\cite{fu_reconstructing_2011} \cite{arrieta-ortiz_experimentally_2015}. 
We estimate TFA based on known regulatory interactions using network component analysis \cite{liao2003network}.
To test the integration of this approach with our fused regression, we assessed the combination of \textit{B. subtilis} datasets, as in Figure \ref{fusedl2-real}a, with the incorporation of TFA estimation. 
We randomly divided the prior known interactions in half, and used half to learn TFA and to generate priors on network structure. 
The remaining interactions were reserved as a gold standard for validation. 
As in previous studies, we observed a marked improvement in network inference when using transcription factor activity (Figure \ref{tfa}). We also obtained AUPR improvement when using fused regression alongside TFA; gains from sharing information across datasets using fused regression were preserved and even enhanced by using TFA.

@@

This is equivalent to the original problem due to the block structure of matrix multiplication. 
In an ordinary regression problem each response variable can be solved independently, and vectorization is unnecessary. 
However, in fused regression, we append additional rows to the design matrix that link entries of the interaction weight matrix associated with different response variables (Figure \ref{cartoon}). 
As a result, these linked response variables must be solved simultaneously through vectorization. 
Two response variables are linked if any of the regulatory weights affecting those genes are linked by a fusion constraint. 
Two response variables must be solved simultaneously if there is any chain of linked response variables connecting them. 
However, every other response variable can be solved separately. 
In biological terms, the regulators of two genes (whether in the same species, or different species) must be solved together if there is a fusion constraint linking those genes' regulators, or if there is a chain of such constraints. 
If the networks for a large number of genes are solved simultaneously, the system of equations can quickly become intractable.

@@

And for steady state conditions
\begin{equation}
\alpha_{i}x_{i}(t_k) = \sum \beta_{i,j}x_{j}(t_k)
\end{equation}
Both kinds of conditions can be included in the set of equations to be solved. We can summarize these equations in matrix form as
\begin{equation}
Y = X \beta 
\end{equation}
where $Y$ is the gene expression matrix, $X$ is the TF expression matrix, and $\beta$ is the regulatory weights we are interested in learning.
We are interested in learning $\beta$, the matrix representation of the gene regulatory network, where the weight in a given position represents the regulatory weight of a TF on a gene. 
Positive weights represent activation, negative weights represent repression, and 0 weights represent the absence of an interaction. The matrix $\beta$ can be solved using linear regression. 
Because there are typically far fewer conditions than possible regressors (TFs), we introduce a ridge regularization constraint with weight $\lambda_R$ and solve
\begin{equation}
\argmin_\beta\vert \vert X\beta - Y_2 \vert \vert ^2 + \lambda_R \vert \vert \beta \vert \vert ^2
\end{equation}
This is similar to the formulation used in the \textit{Inferelator} algorithm, except that we use an $L2$ penalty in place of the more standard $L1$ penalty. Although $L1$ penalties have the theoretical appeal of sparsity - which matches our assumption of sparsity in biological networks - we are primarily interested here in ranking interactions, rather than recovering a specific network. Moreover, $L2$ penalized regression followed by thresholding can be competitive with $L1$ penalized regression for support recovery \cite{} [LOOK UP THIS REF]. Although our approach can be easily modified to use an $L1$ penalty, we did not observe performance gains when doing so, and the version presented here is in some ways conceptually simpler (with the combined penalty function being equivalent to a multivariate Gaussian prior).

@@

\textit{B. subtilis} and \textit{B. anthracis} are distantly related bacterial species with limited gene orthology. Nevertheless we show that network recovery in \textit{B. subtilis} can be improved through the inclusion of expression data from \textit{B. anthracis}. Many previous methods for cross-species network inference operate on the conserved subset of orthologous genes \cite{dillman_comparative_2015}. This assumption may be appropriate with very closely related species, but could not be applied in this domain, where a large fraction (62\% and 67\%) of the \textit{B. subtilis} and \textit{B. anthracis} genomes do not have clear orthologs. Our method, in contrast, can obtain improvements in network inference performance even when the conserved subset of genes is small.

@@

\pagebreak
\begin{figure}[h!]
    \centering
    \includegraphics[scale=.5,trim=0mm 0mm 0mm 0mm,clip]{party3.eps}
    \caption{\label{cartoon} Schematic representation of design matrix construction. Here, the circles and hexagons correspond to different species. Bidirectional arrows represent orthology information and dotted arrows represent putative interactions between TFs and genes. Rectangles under ``Data'' represent TF $\times$ condition matrices of gene expression values in species 1 (top row) and species 2 (bottom row), colored in correspondence with the gene orthology diagram, with bidirectional arrows representing orthology between the two species. Rectangles under ``Weights'' represent gene regulatory interactions in each species, with lines linking coefficients that are fused due to the orthology information shown to the left. Networks associated with genes in each species can be solved independently unless there exist a fusion constraint constraining coefficients of each gene, or a path of such constraints. When a path does exist, these genes must be solved simultaneously. In this example, genes $B$, $C$, and $B'$ must be solved simultaneously; the lower right corner shows a representation of the design matrix necessary to solve the this fused regression problem.
    }
\end{figure}
\pagebreak

\begin{figure}
\begin{center}
  \includegraphics[scale=0.30]{fig1.pdf}
  \caption{\label{fusion} \textbf{A.} Schematic representation of the the generation of fusion constraints from orthology mappings. Dashed arrows indicate potential regulatory interactions, while solid arrows denote orthology. We introduce fusion constraints for pairs of interactions for which both the regulator and regulated gene are orthologs of one another.   In this example, we would introduce a constraint between the $(A, B)$ and $(A', B')$ interactions and the $(A, C)$ and $(A', B')$ interactions. \textbf{B.} 
  In order to demonstrate the utility of fused network inference in combining data, we generate two networks with 10 TFs and 200 genes (75\% sparsity). Mean squared error of the inferred vs. true coefficient matrices for network 1 are plotted as a function of the number of conditions generated for species 1 (x-axis) and the number of conditions generated for species 2 (y-axis). 
  As expected, increasing the number of samples available for the species of interest improves network inference performance. 
  However, because we are fusing to data from a related species, similar gains are observed when increasing the amount of data available in this second species. 
  \textbf{C-F} Show the varying effects of fusion on simulated networks with different levels of conservation. We generate a series of networks with 20 TFs by 200 genes in two species, (50\% sparsity) while varying the fraction of gene orthologies in the simulated networks (Amount fused, shown above each figure). For each network, we evaluated AUPR on one of the species for: all interactions (blue line), interactions with fusion constraints (green line), and interactions without fusion constraints (orange line). At every level of conservation, constrained interactions show the largest benefits of fusion, with the magnitude of the benefit growing with fraction of orthologous genes. 
  When the networks are highly conserved, however, even interactions that are not directly constrained through fusion are recovered more accurately as $\lambda_S$ increases.}
\end{center}
\end{figure}
\pagebreak

@@

\begin{figure}
\begin{center}
  \includegraphics[scale=0.30]{fig3.pdf}
  \caption{\label{adapt-schematic} Adaptive fusion loss function (\textbf{A}) and derivative of loss function (\textbf{B}). 
  \textbf{A.} Adaptive fusion is a quadratic around the origin, begins to taper at $a/2$, and plateaus at $a$. After the plateau, increasing the difference in interaction weight of fused interactions does not further affect the penalty incurred through fusion. As a result, interaction weights in this zone are effectively unfused from one another (the fusion penalty behaves like a constant). 
  \textbf{B.} Shows the derivative of the adaptive fusion penalty, which is used to implement adaptive fusion through local quadratic approximation. The adaptive fusion penalty is modified from SCAD (smoothly clipped absolute deviation) and MCP (minimax concave penalty) functions and like these penalties has a zero derivative far from the origin. }
\end{center}
\end{figure}
\pagebreak

\begin{figure}
\begin{center}
  \includegraphics[scale=0.30]{figure_4_rev1.pdf}
  \caption{\label{adaptivefusion} In order to evaluate the performance of adaptive fusion in the presence of non-conserved interactions, we performed a series of simulations inferring networks given a partially corrupted list of orthology mappings (or, orthology mappings that do not represent functional similarity).
    Networks were generated with 35 TFs by 200 genes, 60\% orthology coverage, 40\% false orthology coverage, and 30 samples per network. \textbf{A.} Top: we plot the interaction weights between pairs of fused interactions in network 1 (x-axis) and network 2 (y-axis) following network inference without fusion ($\lambda_S=0$). True fusion constraints are generated from pairs of true orthologs, while false fusion constraints derive from a pair of orthologs at least one of which is false. When $\lambda_S=0$ -- equivalent to fitting the networks independently -- interactions linked by false constraints are uncorrelated with one another, while interactions linked by true constraints are correlated, reflecting their functional similarity. Below we plot the distribution of differences in weights joined by true and false constraints; true constraints have differences on average closer to zero.\textbf{B.} shows the weights of interactions when networks are fit using fused-L2. In this condition, interactions bound by both true and false fusion constraints are forced to be very similar across the two species. \textbf{C.} With adaptive fusion, many of the false fusion constraints are relaxed. These constraints are no longer forced to be similar, and some of the original structure is restored. \textbf{D.} Shows the effect of fused-L2 and adaptive fusion on the accuracy of network recovery, as measured by MSE between the known simulated network weights and the inferred network weights. When fusion is introduced, recovery of the correctly fused part of the network improves, but the bias induced by fused-L2 degrades performance on the incorrectly fused part of the network. With adaptive fusion, the gains on the conserved part of the network are preserved with significantly less loss of accuracy on the non-conserved part of the network. Error bars are $95\%$ confidence intervals on the squared error of coefficients.}
\end{center}
\end{figure}
